Refinement for open automata

### 2021-05-06 2021-Refinement/report.tex
Paid half of lemma1 proof debt

\(S, V, J\) are arbitrary finite sets.
\end{defi}
The variable names may clash when considering two automata, in that case we suppose that we can still distinguish the variables in the formulas.
% The open transitions will be considered to be a finite set of representative open transitions. Removing this restriction may break some theorems.
The initial valuation may be a partial valuation of the variables, if the undefined variables are set before being read this behave like a normal variable.
If the undefined variable is not set before being read then its value may be any fixed value, the potential use of this construct will be explain in more details in section ???\marginpar{refinement section, hole identical is sufficient for the parametric construct}.
\begin{defi}[Open transition]
An open transition is a tuple \nmm{\OTg} with \(s, s' \in S\) the source and target states, \(\alpha \in \actions\) the produced action, \(J' \subseteq J\) the holes involved, \(\beta_j \in \actions\) the actions of the holes, \(g \in \rformulae\) the guard and \(\psi: V \to \terms\) the new value of the variables.
\end{defi}
An open transition can have many unbound variable, actually an effective transition of the automaton is any well-formed substitution of the unbound variables of the transition minus the automaton variables.
This is used to constrain the infinite branching to be representable, making it possible to reason about.

@@

We can proceed to explain the left enable operator example.
The automata on the left side is \(\OA{\mbrc{L, R}}{L}{\emptyset}{\mbrc{}}{\mbrc{l, r}}{T}\) where transitions in \(T\) are generated by
\begin{align*}
	\OT{L}{L}{x}{\mbrc{l \mapsto x}}{x \neq \delta}{\mbrc{}} &&
	\OT{L}{R}{\tau}{\mbrc{l \mapsto \delta\mpar{x}, r \mapsto \act{acc}\mpar{y}}}{x = y}{\mbrc{}} &&
	\OT{R}{R}{x}{\mbrc{r \mapsto x}}{\top}{\mbrc{}}
\end{align*}
Note that we could have expressed the transition in the middle with \(\mbrc{l \mapsto \delta\mpar{x}, r \mapsto \act{acc}\mpar{x}}\) which would have avoided many transitions with trivially false guards like the one where \(x \mapsto 1, y \mapsto 2\) which has the guard \(1 = 2\).

@@

Finally we can define some utilitary functions:
\begin{defi}[Guard, Out-transition, Transition variables]
Let \(V\) be the variable names of the considered automaton, \(T\) its transitions and \(s\) be one of its states.
\(\fOT{s}\) are called the out-transitions of \(s\).
\(\fIT{s'}\) are called the in-transitions of \(s'\).
\begin{align*}
	\fOT{t} & \defnotation \mset{\OTg \in T}{s = t} &
	\fIT{t} & \defnotation \mset{\OTg \in T}{s' = t} \\
\end{align*}
\begin{gather*}
	\fvars{\OTg} \defnotation \mpar{\fvars{\alpha} \cup \fvars{g} \cup \bigcup_{j \in J'} \fvars{\beta_j} \cup \bigcup_{v \in V} \fvars{\psi\mpar{v}}} \setminus V \\
	\fguard{\OTg} \defnotation g
\end{gather*}
\end{defi}

\begin{exi}[Enable, variable-oriented]
The automata drawn on the right side of the figure is an alternative implementation of the enable operator which is FH-bisimilar to the previous one.
The drawing introduces the double arrows which is used to denote the initial valuation.
The automata drawn is \(\OA{\mbrc{\_}}{\_}{\mbrc{v}}{l^\mbrc{v}}{\mbrc{l, r}}{T}\) where transitions in \(T\) are like previously all the transitions generated by
\begin{align*}
	\OT{\_}{\_}{x}{\mbrc{l \mapsto x}}{v = l \wedge x \neq \delta}{\mbrc{}} &&
	\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \delta\mpar{x}, r \mapsto \act{acc}\mpar{y}}}{v = l \wedge x = y}{\mbrc{v \gets r}} &&
	\OT{\_}{\_}{x}{\mbrc{r \mapsto x}}{v = r}{\mbrc{}}
\end{align*}
Note that in this example \(r\) and \(l\) are a hole names and also closed terms: this is a naming conflict, we do not have \(J \subseteq \rterms\).

The variables of a transition are all the free variables in the expressions minus the auomaton variables.
For instance \(v\) is a variable name for the automaton which is used in the guard.
It is not a transituon variable.
When generating transition without free variables its value is not substituted with a closed term.
For instance \nmm{\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \tau}}{v = l \wedge \tau \neq \delta}{\mbrc{}}} is a transition generated by the second transition and is in \(T\) but not \nmm{\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \tau}}{l = l \wedge \tau \neq \delta}{\mbrc{}}} (result of the forbidden substitution \(v \mapsto l\)).

@@

\section{Semantic and composition of Open Automata}\label{sec:semcompo}
The interpretation of open automata is a partially specified automaton, where the partial specification comes from the holes.
This specification can be completed by composing automata into holes so it is natural to take a look at the composition process.
This was already defined on pNets in previous work \cite{henrio:01299562} but never completely formalised on open automata.
The  definition of composition below is a direct translation of composition on pNets for open automata.
\begin{defi}[Composition of open automata]
The composition of \(A_c \defobject \OA{S_c}{s_{0c}}{V_c}{\sigma_{0c}}{J_c}{T_c}\) in the hole \(k \in J_p\) of \(A_p \defobject \OA{S_p}{s_{0p}}{V_p}{\sigma_{0p}}{J_p}{T_p}\) is
\begin{align*}
	A_p\subst{A_c}{k} \defnotation & \OA{S_p \times S_c}{\mpar{s_{0p}, s_{0c}}}{V_p \uplus V_c}{\sigma_{0p} \uplus \sigma_{0c}}{J_c \uplus J_p \setminus \mbrc{k}}{T} \\
	\text{With } T \defobject & \mset{\OT{\mpar{s_p, s_c}}{\mpar{s'_p, s'_c}}{\alpha_p}{\beta_j^{j \in J'_c \uplus J'_p \setminus \mbrc{k}}}{g_p \wedge g_c \wedge \alpha_c = \beta_k}{\psi_p \uplus \psi_c}}{\OTx{p}{}{}{p} \in T_p, \OTx{c}{}{}{c} \in T_c} \\
	& \cup \mset{\OT{\mpar{s_p, s_c}}{\mpar{s'_p, s_c}}{\alpha_p}{\beta_j^{j \in J'_p}}{g_p}{\psi_p}}{\OTx{p}{}{}{p}, k \notin J', s_c \in S_c}
\end{align*}
Similarly to expression substitution, the parallel composition is noted \(A\psubst{A_j^{j \in J}}\).
\end{defi}
The actions emitted when \(A_c\) makes a transition is sychronised with the action of the hole \(k\) in transitions of \(A_p\) which have it as a hole actions (first transition set, \(\alpha_c = \beta_k\)).
No transition of \(A_c\) is performed when a transition that do not refer to the hole \(k\) is performed in \(A_p\) (second transition set, \(k \notin J'\)).
The composition may look like a handshake, with both automata running in parallel (product of states and joint variables) but it is actually not symmetric.
\begin{exi}
This example is derived and adapted from a traffic light controller in a collection of examples for pNets.
It is supposed to be a single traffic light.

@@

\begin{figure}
\centering
\input{Traffic_Lights_Spec.tex}
\caption{The specification of a traffic light system}
\label{fig:tls}
\end{figure}
\begin{figure}
\centering
\input{Traffic_Lights_Controller.tex}
\vrule
\input{Traffic_Lights_Register.tex}
\caption{On the left: The controller agent; On the right: The counter agent}
\label{fig:tlh}
\end{figure}
The component we choose to compose in the holes are in Figure \ref{fig:tlh}.
On the left is the controller agent which will be composed in the hole \(ctl\).
Its role is to decide the duration before switching the lights.
This implementation is simply a constant time for each color.

@@

The automaton on Figure \ref{fig:tlf} is bisimilar to the composition of the specification on Figure \ref{fig:tls} and the agents on Figure \ref{fig:tlh}.
The result of the composition has 36 states (\(3 \times 2 \times 6\)) but only 6 are reachable from the initial configuration.
The transitions obtained by composition were simplified to give the following (where \(n, n+1\) stands for numbers between 1 and 6 and Xxx is either Red, Yellow or Green):
\begin{itemize}
\item \nmm{\OT{XnS}{X(n+1)C}{\tau}{\mbrc{}}{\top}{\mbrc{t \gets k, c \gets 0}}} which is the composition of \nmm{\OT{X}{X}{\tau}{\mbrc{ctl \mapsto \theta\mpar{x}, cnt \mapsto \act{set}\mpar{x}}}{\top}{\mbrc{}}} from the specification with \nmm{\OT{S}{C}{\act{set}\mpar{x}}{\mbrc{}}{\top}{\mbrc{t \gets x, c \gets 0}}} from the counter and \nmm{\OT{n}{n+1}{\theta\mpar{k}}{\mbrc{}}{\top}{\mbrc{}}} from the controller.
\item \nmm{\OT{XnC}{XnC}{\act{tick}}{\mbrc{}}{c < t}{\mbrc{c \gets c + 1}}} which is the composition of \nmm{\OT{X}{X}{\act{tick}}{\mbrc{cnt \mapsto \act{tick}}}{\top}{\mbrc{}}} from the specification with \nmm{\OT{C}{C}{\act{tick}}{\mbrc{}}{c < t}{\mbrc{c \gets c + 1}}} from the counter.
\item \nmm{\OT{YnC}{X(n+1)S}{\act{TurnXxx}}{\mbrc{}}{c = t}{\mbrc{}}} which is the composition of \nmm{\OT{Y}{X}{\act{TurnXxx}}{\mbrc{cnt \mapsto \act{over}\mpar{x}, ctl \mapsto \delta\mpar{x}}}{\top}{\mbrc{}}} from the specification with \nmm{\OT{C}{S}{\act{over}\mpar{c}}{\mbrc{}}{c = t}{\mbrc{}}} from the counter and \nmm{\OT{n}{n+1}{\delta\mpar{x}}{\mbrc{}}{\top}{\mbrc{}}} from the controller.
\end{itemize}

\begin{figure}
\centering
\input{Traffic_Lights_Full.tex}
\caption{The full traffic lights system}
\label{fig:tlf}
\end{figure}
To illustrate composition, simplification and value passing, the \(\tau\) transition from the state \(R\) will be examined.
The transition in the specification is \nmm{\OT{R}{R}{\tau}{\mbrc{cnt \mapsto \act{set}\mpar{x}, ctl \mapsto \theta\mpar{x}}}{\top}{\mbrc{}}}.
The composition can produce 12 states containing the \(R\) state, these states are in \(\mbrc{R} \times \mdbrk{1; 6} \times \mbrc{S, C}\), however only \(R1S\) and \(R2C\) are reachable so we will only consider \(\tau\) transitions from these two.
The holes \(cnt\) and \(ctl\) are both involved in the \(\tau\) transition and filled with automata so the composed automata must synchronise with a transition.

@@

The relations will have to be at least preorders, simulations and \Ludo{introduction of deadlocks}{define more precisely introduction of deadlocks or refer to a future definition} (state where no transition is possible) to be called refinement simulation.
On top of that weak refinement are refinement where some actions, typically one named \(\tau\), are considered not observable and can occur in the implementation while not having to be matched in the specification.
In the weak refinement the relations have to forbid the introduction of livelocks (only a sequence of \(\tau\) possible) and respect \(\tau\)-stuttering \marginpar{cite many things arround here}.
\Ludo{to add}{I believe we want more arguments on why we want some properties, at least deadlock and livelock related, perhaps also weak vs strong. How did you decide of the list of useful properties / when ais such or such property useful?}

While being preorder is simple to express for a relation, being a simulation, do not introduce new deadlocks/livelocks and \(\tau\)-stuttering are properties about the states of the related automata, therefore they are more complex to express.
We will need a preliminary concept before defining them.
\begin{defi}[Relation under predicate]
A relation under predicate between states of two automata \(\OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\) and \(\OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\) is a function \(R: S_1 \times S_2 \to \rformulae[V_1 \uplus V_2]\).

Two states \(s_1 \in S_1, s_2 \in S_2\) are related under a \Ludo{?}{relation is mission: under a relation $R$ for a valuation ... ?} valuation \(\sigma: V_1 \uplus V_2 \to \values\) iff \(\sigma \vdash R\mpar{s_1, s_2}\).
For a relation \(\wrel{}{}{}\) which uses a relation under predicate to prove that two element are related \(\forall x \, x', x \wrel{}{}{} x' \iff \exists R, P\mpar{R}\), the fact that there is a valid witness \(R\) is noted \(\wrel{x}{x'}{R}\).
\end{defi}
Now the properties about states of automata can be adapted to open automata.
\begin{defi}[Simulation on open automata]
A relation \(\leq\) is a simulation if for any two related automata \(\wrel{A_1}{A_2}{R}\) where \(A_1 \defobject \OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\), \(A_2 \defobject \OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\) and \(R: S_1 \times S_2 \to \rformulae[V_1 \uplus V_2]\), \(R\) satisfies both
\defitem Initial states are related under initial valuations: \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\);
\defitem From related states, all out-transitions from \(A_1\) can be simulated in \(A_2\) and their target states are \Ludo{related}{I am a bit surprised that $J_1$ and $J_2$ are unrelated and we only need to take only the intersection TBD}:
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}, \forall \sigma: V_1 \uplus V_2 \uplus \fvars{t_1} \to \values, \\
	\mpar{\sigma \vdash R\mpar{s_1, s_2} \wedge g_1} \implies \bigsymb{\exists} t_2 \defobject \OTx{2}{}{2}{2} \in \fOT{s_2}, \exists \nu: \fvars{t_2} \to \values, \\
	\sigma \uplus \nu \vdash \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \wedge g_2 \wedge R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}
\end{multline*}
\end{defi}
The last formula means that for every pair of related states under valuation of the automaton variables and every possible transitions (\(\vdash g_1\)) from the first automaton, there is a possible transition (\(\vdash g_2\)) such that the produced action matches (\(\alpha_1 = \alpha_2\)), the holes actions from same hole names (\(J'_1 \cap J'_2\)) match (\(\beta_{1j} = \beta_{2j}\)) and the target states are related after variable update.
This is a natural extension of the notion of simulation on LTS, which is the same definition without variables and holes and guards.
It will serve as a watchdog for any relation that will be defined with ``simulation" in its name, except weak-simulation where the transition from the first automaton that produce \(\tau\) actions are excluded.

\begin{defi}[Deadlock reduction, intuitive definition]
A simulation \(\leq\) is deadlock reducing if in all the \(R: S_1 \times S_2 \to \rformulae[V_1 \uplus V_2]\) such that \(\wrel{A_1}{A_2}{R}\), there is one that also satisfies
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \forall \sigma: V \to \values, \mpar{\sigma \vdash R\mpar{s_1, s_2}} \implies \\
	\mpar{\everymath{\displaystyle}\begin{array}{c}
		\bigsymb{\exists} t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}, t_2 \defobject \OTx{2}{}{2}{2} \in \fOT{s_2}, \exists \nu: \fvars{t_1} \uplus \fvars{t_2} \to \values, \\
		\sigma \uplus \nu \vdash g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \wedge R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}
	\end{array}} \\
	\vee \forall t_2 \in \fOT{s_2}, \forall \nu: \fvars{t_2} \to \values, \sigma \uplus \nu \nvdash \fguard{t_2}
\end{multline*}
\end{defi}
This definition \Ludo{extends the requirements of a simulation}{je suis pas d accord avec cette affirmation, ou tout du moins c est pas evident} by requiring from the relation that on related states, either there is a pair of possible matching transition or there was no out-transition possible in the specification.
This requirement effectively prevents the presence of deadlock state in the first automaton if there is a related non-deadlock state in the second.

@@

``Deadlock reduction" conflicts with ``composition completeness" by disallowing some unwanted cases where an automaton in a hole cannot produce any action that any out-transitions expects.
In particular filling a hole with the deadlock automaton (only one state without out-transitions) is not considered a refinement for a \Ludo{no deadlock reducing simulation refinement}{define or use the name of the definition}.
A way to solve this \Ludo{issue}{why is this an issue? can be solved by stating which properties are desirable for which reason above} is to characterise a composition that do not introduce deadlocks, which will be used instead of the composition introduced earlier.
\begin{defi}[Reachability]
For any open automata \(A \defobject \OA{S}{s_0}{V}{\sigma_0}{J}{T}\), a reachability predicate \(\reach{A}: S \to \rformulae[V]\) is a predicate on states such that
\defitem \(\sigma_0 \vdash \reach{A}\mpar{s_0}\).
\defitem \nmm{\bigsymb{\forall} t \defobject \OTg \in T, \forall \nu: \fvars{t} \to \values, \nu \vdash \reach{A}\mpar{s} \wedge g \implies \reach{A}\mpar{s'}\psubst{\psi}}.
\end{defi}

>>>>

\(S, V, J\) are arbitrary finite sets.
\end{defi}
The variable names may clash when considering two automata, in that case we suppose that we can still distinguish the variables in the formulas.
% The open transitions will be considered to be a finite set of representative open transitions. Removing this restriction may break some theorems.
The initial valuation may be a partial valuation of the variables, if the undefined variables are set before being read this behave like a normal variable.
If the undefined variable is not set before being read then its value may be any fixed value, the potential use of this construct will be explain in more details in section ???\marginpar{refinement section, hole identical is sufficient for the parametric construct}.
\begin{defi}[Open transition]
An open transition is a tuple \nmm{\OTg} with \(s, s' \in S\) the source and target states, \(\alpha \in \actions\) the produced action, \(J' \subseteq J\) the holes involved, \(\beta_j \in \actions\) the actions of the holes, \(g \in \rformulae\) the guard and \(\psi: V \to \terms\) the new value of the variables.
\end{defi}
An open transition can have many unbound variable, actually an effective transition of the automaton is any well-formed substitution of the unbound variables of the transition minus the automaton variables.
% This is used to constrain the infinite branching to be representable, making it possible to reason about.

@@

We can proceed to explain the left enable operator example.
The automata on the left side is \(\OA{\mbrc{L, R}}{L}{\emptyset}{\mbrc{}}{\mbrc{l, r}}{T}\) where transitions in \(T\) are:
\begin{align*}
	\OT{L}{L}{x}{\mbrc{l \mapsto x}}{x \neq \delta}{\mbrc{}} &&
	\OT{L}{R}{\tau}{\mbrc{l \mapsto \delta\mpar{x}, r \mapsto \act{acc}\mpar{y}}}{x = y}{\mbrc{}} &&
	\OT{R}{R}{x}{\mbrc{r \mapsto x}}{\top}{\mbrc{}}
\end{align*}
Note that we could have expressed the transition in the middle with \(\mbrc{l \mapsto \delta\mpar{x}, r \mapsto \act{acc}\mpar{x}}\) which would have avoided many transitions with trivially false guards like the one where \(x \mapsto 1, y \mapsto 2\) which has the guard \(1 = 2\).

@@

Finally we can define some utilitary functions:
\begin{defi}[Guard, Out-transition, Transition variables]
Let \(V\) be the variable names of the considered automaton, \(T\) its transitions and \(s\) be one of its states.
\(\fOT{s}\) are called the out-transitions of \(s\).
\(\fIT{s'}\) are called the in-transitions of \(s'\).
\begin{align*}
	\fOT{t} & \defnotation \mset{\OTg \in T}{s = t} &
	\fIT{t} & \defnotation \mset{\OTg \in T}{s' = t} \\
\end{align*}
\vspace{-1cm}
\begin{gather*}
	\fguard{\OTg} \defnotation g \\
	\fvars{\OTg} \defnotation \mpar{\fvars{\alpha} \cup \fvars{g} \cup \bigcup_{j \in J'} \fvars{\beta_j} \cup \bigcup_{v \in V} \fvars{\psi\mpar{v}}} \setminus V
\end{gather*}
\end{defi}

\begin{exi}[Enable, variable-oriented]
The automata drawn on the right side of the figure is an alternative implementation of the enable operator which is FH-bisimilar to the previous one.
The drawing introduces the double arrows which is used to denote the initial valuation.
The automata drawn is \(\OA{\mbrc{\_}}{\_}{\mbrc{v}}{l^\mbrc{v}}{\mbrc{l, r}}{T}\) where transitions in \(T\) are:
\begin{align*}
	\OT{\_}{\_}{x}{\mbrc{l \mapsto x}}{v = l \wedge x \neq \delta}{\mbrc{}} &&
	\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \delta\mpar{x}, r \mapsto \act{acc}\mpar{y}}}{v = l \wedge x = y}{\mbrc{v \gets r}} &&
	\OT{\_}{\_}{x}{\mbrc{r \mapsto x}}{v = r}{\mbrc{}}
\end{align*}
Note that in this example \(r\) and \(l\) are a hole names and also closed terms: this is a naming conflict, we do not have \(J \subseteq \rterms\).

The variables of a transition are all the free variables in the expressions minus the automaton variables.
For instance \(v\) is a variable name for the automaton which is used in the guard.
It is not a transituon variable.
When generating transition without free variables its value is not substituted with a closed term.
For instance \nmm{\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \tau}}{v = l \wedge \tau \neq \delta}{\mbrc{}}} is a transition generated by the second transition and is in \(T\) but not \nmm{\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \tau}}{l = l \wedge \tau \neq \delta}{\mbrc{}}} (result of the forbidden substitution \(v \mapsto l\)).

@@

\section{Semantic and composition of Open Automata}\label{sec:semcompo}
The interpretation of open automata is a partially specified automaton, where the partial specification comes from the holes.
This specification can be completed by composing automata into holes so it is natural to take a look at the composition process.
This was already defined on pNets in previous work \cite{henrio:01299562} but never completely formalised on open automata.
The  definition of composition below is a direct translation of composition on pNets for open automata.
\begin{defi}[Composition of open automata]
The composition of \(A_c \defobject \OA{S_c}{s_{0c}}{V_c}{\sigma_{0c}}{J_c}{T_c}\) in the hole \(k \in J_p\) of \(A_p \defobject \OA{S_p}{s_{0p}}{V_p}{\sigma_{0p}}{J_p}{T_p}\) is
\begin{align*}
	A_p\subst{A_c}{k} \defnotation & \OA{S_p \times S_c}{\mpar{s_{0p}, s_{0c}}}{V_p \uplus V_c}{\sigma_{0p} \uplus \sigma_{0c}}{J_c \uplus J_p \setminus \mbrc{k}}{T} \\
	\text{With } T \defobject & \mset{\OT{\mpar{s_p, s_c}}{\mpar{s'_p, s'_c}}{\alpha_p}{\beta_j^{j \in J'_c \uplus J'_p \setminus \mbrc{k}}}{g_p \wedge g_c \wedge \alpha_c = \beta_k}{\psi_p \uplus \psi_c}}{\OTx{p}{}{}{p} \in T_p, \OTx{c}{}{}{c} \in T_c} \\
	& \cup \mset{\OT{\mpar{s_p, s_c}}{\mpar{s'_p, s_c}}{\alpha_p}{\beta_j^{j \in J'_p}}{g_p}{\psi_p}}{\OTx{p}{}{}{p}, k \notin J', s_c \in S_c}
\end{align*}
Similarly to expression substitution, the parallel composition is noted \(A\psubst{A_j^{j \in J}}\).
\end{defi}
The actions emitted when \(A_c\) makes a transition is sychronised with the action of the hole \(k\) in transitions of \(A_p\) which have it as a hole actions (first transition set, \(\alpha_c = \beta_k\)).
No transition of \(A_c\) is performed when a transition that do not refer to the hole \(k\) is performed in \(A_p\) (second transition set, \(k \notin J'\)).
The composition may look like a handshake, with both automata running in parallel (product of states and joint variables) but it is actually not symmetric.
\begin{exi}
\begin{figure}
\centering
\input{Traffic_Lights_Spec.tex}
\caption{The specification of a traffic light system}
\label{fig:tls}
\end{figure}
\begin{figure}
\centering
\input{Traffic_Lights_Controller.tex}
\vrule
\input{Traffic_Lights_Register.tex}
\caption{On the left: The controller agent; On the right: The counter agent}
\label{fig:tlh}
\end{figure}
This example is derived and adapted from a traffic light controller in a collection of examples for pNets.
It is supposed to be a single traffic light.

@@

The component we choose to compose in the holes are in Figure \ref{fig:tlh}.
On the left is the controller agent which will be composed in the hole \(ctl\).
Its role is to decide the duration before switching the lights.
This implementation is simply a constant time for each color.

@@

\begin{figure}
\centering
\input{Traffic_Lights_Full.tex}
\caption{The full traffic lights system}
\label{fig:tlf}
\end{figure}
The automaton on Figure \ref{fig:tlf} is bisimilar to the composition of the specification on Figure \ref{fig:tls} and the agents on Figure \ref{fig:tlh}.
The result of the composition has 36 states (\(3 \times 2 \times 6\)) but only 6 are reachable from the initial configuration.
The transitions obtained by composition were simplified to give the following (where \(n, n+1\) stands for numbers between 1 and 6 and Xxx is either Red, Yellow or Green):
\begin{itemize}
\item \nmm{\OT{XnS}{X(n+1)C}{\tau}{\mbrc{}}{\top}{\mbrc{t \gets k, c \gets 0}}} which is the composition of \nmm{\OT{X}{X}{\tau}{\mbrc{ctl \mapsto \theta\mpar{x}, cnt \mapsto \act{set}\mpar{x}}}{\top}{\mbrc{}}} from the specification with \nmm{\OT{S}{C}{\act{set}\mpar{x}}{\mbrc{}}{\top}{\mbrc{t \gets x, c \gets 0}}} from the counter and \nmm{\OT{n}{n+1}{\theta\mpar{k}}{\mbrc{}}{\top}{\mbrc{}}} from the controller.
\item \nmm{\OT{XnC}{XnC}{\act{tick}}{\mbrc{}}{c < t}{\mbrc{c \gets c + 1}}} which is the composition of \nmm{\OT{X}{X}{\act{tick}}{\mbrc{cnt \mapsto \act{tick}}}{\top}{\mbrc{}}} from the specification with \nmm{\OT{C}{C}{\act{tick}}{\mbrc{}}{c < t}{\mbrc{c \gets c + 1}}} from the counter.
\item \nmm{\OT{YnC}{X(n+1)S}{\act{TurnXxx}}{\mbrc{}}{c = t}{\mbrc{}}} which is the composition of \nmm{\OT{Y}{X}{\act{TurnXxx}}{\mbrc{cnt \mapsto \act{over}\mpar{x}, ctl \mapsto \delta\mpar{x}}}{\top}{\mbrc{}}} from the specification with \nmm{\OT{C}{S}{\act{over}\mpar{c}}{\mbrc{}}{c = t}{\mbrc{}}} from the counter and \nmm{\OT{n}{n+1}{\delta\mpar{x}}{\mbrc{}}{\top}{\mbrc{}}} from the controller.
\end{itemize}

To illustrate composition, simplification and value passing, the \(\tau\) transition from the state \(R\) will be examined.
The transition in the specification is \nmm{\OT{R}{R}{\tau}{\mbrc{cnt \mapsto \act{set}\mpar{x}, ctl \mapsto \theta\mpar{x}}}{\top}{\mbrc{}}}.
The composition can produce 12 states containing the \(R\) state, these states are in \(\mbrc{R} \times \mdbrk{1; 6} \times \mbrc{S, C}\), however only \(R1S\) and \(R2C\) are reachable so we will only consider \(\tau\) transitions from these two.
The holes \(cnt\) and \(ctl\) are both involved in the \(\tau\) transition and filled with automata so the composed automata must synchronise with a transition.

@@

The relations will have to be at least preorders, simulations and forbid the introduction of deadlocks (state where no transition is possible) to be called refinement simulation.
On top of that weak refinement are refinement where some actions, typically one named \(\tau\), are considered not observable and can occur in the implementation while not having to be matched in the specification.
In the weak refinement the relations have to forbid the introduction of livelocks (only a sequence of \(\tau\) possible) and respect \(\tau\)-stuttering \marginpar{cite many things arround here}.
While being preorder is simple to express for a relation, being a simulation, do not introduce new deadlocks/livelocks and \(\tau\)-stuttering are properties about the states of the related automata, therefore they are more complex to express.
We will need a preliminary concept before defining them.
\begin{defi}[Relation under predicate]
A relation under predicate between states of two automata \(\OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\) and \(\OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\) is a function \(R: S_1 \times S_2 \to \rformulae[V_1 \uplus V_2]\).

Two states \(s_1 \in S_1, s_2 \in S_2\) are related under a valuation \(\sigma: V_1 \uplus V_2 \to \values\) iff \(\sigma \vdash R\mpar{s_1, s_2}\).
For a relation \(\wrel{}{}{}\) which uses a relation under predicate to prove that two element are related \(\forall x \, x', x \wrel{}{}{} x' \iff \exists R, P\mpar{R}\), the fact that there is a valid witness \(R\) is noted \(\wrel{x}{x'}{R}\).
\end{defi}
Now the properties about states of automata can be adapted to open automata.
\begin{defi}[Simulation on open automata]
A relation \(\leq\) is a simulation if for any related automata \(\wrel{A_1}{A_2}{R}\) where \(A_1 \defobject \OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\), \(A_2 \defobject \OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\) and \(R: S_1 \times S_2 \to \rformulae[V_1 \uplus V_2]\), \(R\) satisfies both
\defitem Initial states are related under initial valuations: \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\);
\defitem From related states, all out-transitions from \(A_1\) can be simulated in \(A_2\) and their target states are related:
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}, \forall \sigma: V_1 \uplus V_2 \uplus \fvars{t_1} \to \values, \\
	\mpar{\sigma \vdash R\mpar{s_1, s_2} \wedge g_1} \implies \bigsymb{\exists} t_2 \defobject \OTx{2}{}{2}{2} \in \fOT{s_2}, \exists \nu: \fvars{t_2} \to \values, \\
	\sigma \uplus \nu \vdash \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \wedge g_2 \wedge R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}
\end{multline*}
\end{defi}
The last formula means that for every pair of related states under valuation of the automaton variables and every possible transitions (\(\vdash g_1\)) from the first automaton, there is a possible transition (\(\vdash g_2\)) such that the produced action matches (\(\alpha_1 = \alpha_2\)), the holes actions from same hole names (\(J'_1 \cap J'_2\)) match (\(\beta_{1j} = \beta_{2j}\)) and the target states are related after variable update.
This is a natural extenion of the notion of simulation on LTS, which is the same definition without variables and holes and guards.
It will serve as a watchdog for any relation that will be defined with ``simulation" in its name, except weak-simulation where the transition from the first automaton that produce \(\tau\) actions are excluded.

\begin{defi}[Deadlock reduction, intuitive definition]
A simulation \(\leq\) is deadlock reducing if in all the \(R: S_1 \times S_2 \to \rformulae[V_1 \uplus V_2]\) such that \(\wrel{A_1}{A_2}{R}\), there is one that also satisfies
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \forall \sigma: V \to \values, \mpar{\sigma \vdash R\mpar{s_1, s_2}} \implies \\
	\mpar{\everymath{\displaystyle}\begin{array}{c}
		\bigsymb{\exists} t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}, t_2 \defobject \OTx{2}{}{2}{2} \in \fOT{s_2}, \exists \nu: \fvars{t_1} \uplus \fvars{t_2} \to \values, \\
		\sigma \uplus \nu \vdash g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \wedge R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}
	\end{array}} \\
	\vee \forall t_2 \in \fOT{s_2}, \forall \nu: \fvars{t_2} \to \values, \sigma \uplus \nu \nvdash \fguard{t_2}
\end{multline*}
\end{defi}
This definition extends the requirements of a simulation by requiring from the relation that on related states, either there is a pair of possible matching transition or there was no out-transition possible in the specification.
This requirement effectively prevents the presence of deadlock state in the first automaton if there is a related non-deadlock state in the second.

@@

``Deadlock reduction" conflicts with ``composition completeness" by disallowing some unwanted cases where an automaton in a hole cannot produce any action that any out-transitions expects.
In particular filling a hole with the deadlock automaton (only one state without out-transitions) is not considered a refinement for a no deadlock reducing simulation refinement.
A way to solve this issue is to caracterise a composition that do not introduce deadlocks, which will be used instead of the composition introduced earlier.
\begin{defi}[Reachability]
For any open automata \(A \defobject \OA{S}{s_0}{V}{\sigma_0}{J}{T}\), a reachability predicate \(\reach{A}: S \to \rformulae[V]\) is a predicate on states such that
\defitem \(\sigma_0 \vdash \reach{A}\mpar{s_0}\).
\defitem \nmm{\bigsymb{\forall} t \defobject \OTg \in T, \forall \nu: \fvars{t} \to \values, \nu \vdash \reach{A}\mpar{s} \wedge g \implies \reach{A}\mpar{s'}\psubst{\psi}}.
\end{defi}
For the reader not used to simulations and bisimulations, this definition of a predicate can seem strange.
The reachability predicate is used to caracterise the states and valuations that are reachable in a run of an automaton.
In fact the role of the predicate is to caracterise potentially reachable configurations without having to caracterise the fact that the state and valuation are the result of a valid path in an automaton.

To do that we impose that the initial state is reachable in the initial valuation and that reachability is preserved by taking valid transitions.
This effectively makes reachability account for all paths and potentially over-approximate the reachable configurations.
However the exact rechability may not be reprensentable in the formulae, hence the need of potentially over-approximating.
